---
title: "Introduction to pinfo201"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pinfo201}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pinfo201)
```

# Summary 
This package enables instructors to efficiently create practice sets for
students who are learning to code in R. It is intended for learners who are just
beginning.

Practice sets include coding prompts, answers, and, optionally, hints. Learners
write solutions in R to the coding prompts and receive feedback. In addition,
this package provides functions, intended for instructors, for automatically
grading student work.

This package installs an Addin menu into RStudio, which learners use to select
practice sets and have their work checked. Instructors are given a set of
administrative functions for creating practice sets and for grading student
work.

Practice sets are specified with a small number of markup tags, which are used
to wrap R code.

# Practice Sets for Learning to Code in R
The key idea is that an instructor creates a practice set, which comprises
prompts, expected answers, and hints. Once the package is installed,
practice sets are loaded into RStudio and can be controlled with an Addins
menu. The Addin menu includes these options:

  * 1-Begin Practice
  * 2-Check Answers
  * 3-Show Prompts
  * 4-Show Answers

Learners select from the available practice sets (_1. Begin Practice_). They write code for one or more of the prompts. When ready, learners check their code (_2. Check Answers_). If the learner's answer
differs from the expected answer, hints, if available, are shown.

The checking process, works with two _variables spaces_: Learner's and expected 
code environments. Basically, the process works as 
follows: 

   1. Execute the learner's code (learner's code environment)
   1. Execute the expected code (expected code environment)
   1. Compare the values of each variable in the learner's code environment with the corresponding values of the variables in the expected code environment
   
**Key idea**: The evaluation process is done by comparing the computed values _after the entire script has been executed_. 

There are two extensions to this basic process: 

  1. **Pre-sets**. A set of variables can be initialized prior to executing the code. Think of these 
  variables as global static variables. They are available in both the 
  learner's and expected code environments. Pre-sets are useful when, for example, a large dataframe needs to be loaded and then processed.
  
  2. **Copy variables**. Prior to executing the expected code those variables that have been marked as `copy variables` are copied to the expected code space. This allows the 
  checking process to be dependent upon variables that are initialized by 
  the learner.

# Practice Sets Mark-up
Instructors write practice sets in .R files, with tagged comments. Here is an example:

```
#' @version ps-1
#' @short PS-Example
#' @title Example practice set
#' @descr
#' This file illustrates the essentials of specifying a practice set. It
#' shows the different kinds of prompts that can be created for learners.
#' @end
#' @initial-vars
library(stats)
X <- c(1,2,3)
#' @end

#' @id -
#' @msg
#' A practice set that illustrates the basics.
#' @end

#' @id a
#' @msg Add ten, nine, and eight together. Assign the result to `sum1`.
#' @code
sum1 <- 10 + 9 + 8
#' @end
#' @hints
#' Do you use the math plus operator (+)?
#' Do you use the assignment operator (<-)?
#' Is the expected variable name (sum1) used?
#' @end

#' @id b
#' @msg
#' Create a variable `hometown` that stores the city in which you were born.
#' @end
#' @check list(re="^[a-zA-Z\\s\\.]*$")
#' @code
hometown <- "St. Louis"
#' @end

#' @id c
#' @msg Add 10 to each of the elements of vector `X`.
#' @code
v1 <- X + 10
#' @end

#' @id d
#' @msg
#' Create 100 random numbers between 40000 and 50000. Round the numbers
#' to two decimal places.
#' @end
#' @cp-var salaries_2017
#' @code
salaries_2017 <- round(runif(100, 40000, 50000),2)
#' @end

#' @id -
#' @msg Working with functions.

#' @id e
#' @msg Write a function, named `what_is_pi`, which returns pi (3.1415).
#' @var what_is_pi
#' @code
what_is_pi <- function() {pi}
#' @end

#' @id f
#' @msg Write a function, named `squared(x)`, which squares a number.
#' @var squared
#' @check list(arg1 = c(1, 2, 3, 0, -1, -2, -3, NA))
#' @code
squared <- function(x) {
   t <- x^2
   return(t)
}
#' @end

#' @id g
#' @msg
#' Define a function, named `imperial_to_metric`, that takes in two arguments: a
#' number of feet and a number of inches. The function should return the
#' equivalent length in meters.
#' @end
#' @check list(arg1 = c(4, 5, 100, 0, NA), arg2 = c(0, 1, 1.5, 12.0, 24))
#' @code
imperial_to_metric <- function(feet, inches) {
   total_inches <- feet * 12 + inches
   meters <- total_inches * 0.0254
   return(meters)
}
#' @end
```
## About a practice sets 
Practice sets are uniquely identified a short name (`@short`). Short IDs should
be valid file names without an `.R` extension.

In addition each practice set has a title (`@title`) and a brief description
(`@descr`). These two tags are optional.

Keep mind that learners will see short IDs, titles, and descriptions.
Accordingly, choose clear and usable language for these tags.

Finally, all practice sets should begin with the version tag (`@version`), which
simply indicates the mark-up version for process practice set files.

```
#' @version ps-1
#' @short PS-Example
#' @title Example practice set
#' @descr
#' This file illustrates the essentials of specifying a practice set. It
#' shows the different kinds of prompts that can be created for learners.
#' @end
```

## Initial, pre-set variables
The `@initial-vars` tag is used to, optionally, initialize one or more
variables. The idea here is that prompts can require learners to use those
variables. As can be seen in the example, specific R libraries can also be
loaded.

```
#' @initial-vars
library(stats)
X <- c(1,2,3)
random_numbers <- round(runif(1000000, 0, 1),4)
#' @end
```

## Coding prompts 
The prompts comprise an ID (`@id`), a  prompt (`@msg`), the expected answer
(`@code`), and a list of hints (`@hints`). 

If an `@id` is a question mark ("?"), when the practice set is loaded into R,
IDs will automatically be generated, which greatly simplifies the development of
practice sets, since during development you do not need to reorder ids when 
new prompts are inserted.

```
#' @id ?
#' @msg Add 10 to each of the elements of vector `X`.
#' @code
v1 <- X + 10
#' @end
```

Students see the prompts as R comments, so they are formatted exactly as they
appear between the `@msg` and `@end` tags.

```
#' @id d
#' @msg
#' Create 100 random numbers between 40000 and 50000. Round the numbers
#' to two decimal places.
#' @end
```

## Message prompts 
If an `@id` is a dash ("-"), a message is formatted. This is a way to guide students through the prompts and provide clarifications.  Message prompts are not check for correctness.

```
#' @id -
#' @msg Working with functions. Next you will practice two things: 
#'    1. Blah, blah, blah ..
#'    2. Wow, wow, wow ... 
#' @end
```

Message prompts can have a code block, indicating what students are expected to code. These 
code blocks, while not checked for correctness, are evaluated so that later prompts can 
depend on them. 

```
#' @id -
#' @msg Working with functions. Next you will practice two things: 
#'    1. Blah, blah, blah ..
#'    2. Wow, wow, wow ... 
#' @end
#' @code
v1 <- X + 10
#' @end
```

## Ordinary variables 
When a practice set is installed, the code block is scanned and the
left-hand-side variable of the _last_ assignment statement is extracted. It is
this variable that is to be checked.

Sometimes it is necessary or convenient to explicitly specify the variable with
the variable tag (`@var`). For example, in the following the variable to be
checked is `squared` not `t`.

For most prompts, that `@var` tag is not needed; instead, let the system 
automatically extract the variable name.

(Note: If you specify a variable name with `@var` and the variable name is 
not in the code block an internal error will be reported at runtime.)

```
#' @id f
#' @msg Write a function, named `squared(x)`, which squares a number.
#' @var squared
#' @check list(arg1 = c(1, 2, 3, 0, -1, -2, -3, NA))
#' @code
squared <- function(x) {
 t <- x^2
 return(t)
}
t <- squared(4)
#' @end
```

In this example, the variable `v1` is extracted. (Note: The variable `X` was
initialized earlier in the practice set.)

```
#' @id c
#' @msg Add 11 to each of the elements of vector `X`.
#' @code
t <- 1
v1 <- X + 10 + t
#' @end
```

## Copying variables
Sometimes it is necessary to copy the value of a variable from the learner's
code environment to the expected code environment. In this example, the
learner's code creates a variable, `salaries_2017`, containing a vector of
random numbers. 

The tag `@cp-var` causes this variable to be copied to the expected code
environment. This allows the checking process to use the values computed by the
learner's code environment.

```
#' @id d
#' @msg
#' Create 100 random numbers between 40000 and 50000. Round the numbers
#' to two decimal places.
#' @end
#' @cp-var salaries_2017
#' @code
salaries_2017 <- round(runif(100, 40000, 50000),2)
#' @end
```

## Checking functions
The `@checks` tag can be used to check the implementation of a
function. Currently, functions with zero, one, or two arguments can be 
automatically checked. For functions with three or more parameters, 
prompt-specific callback functions can be implemented (see below).

In this example, each value in the `arg1` vector is used to check 
the function. 

```
#' @id f
#' @msg Write a function, named `squared(x)`, which squares a number.
#' @var squared
#' @check list(arg1 = c(1, 2, 3, 0, -1, -2, -3, NA))
#' @code
squared <- function(x) {
   t <- x^2
   return(t)
}
#' @end
```

For functions with two parameters, the function is tested on all combinations of
inputs, `arg1` and `arg2`.

```
#' @id g
#' @msg
#' Define a function, named `imperial_to_metric`, that takes in two arguments: a
#' number of feet and a number of inches. The function should return the
#' equivalent length in meters.
#' @end
#' @check list(arg1 = c(4, 5, 100, 0, NA), arg2 = c(0, 1, 1.5, 12.0, 24))
#' @code
imperial_to_metric <- function(feet, inches) {
   total_inches <- feet * 12 + inches
   meters <- total_inches * 0.0254
   return(meters)
}
```

More complex function checking, including checking functions with three or more
arguments, is possible by implementing a prompt-specific callback function (see
below).

# What do learners see?:

The above markup is used to produce the following practice set, a `.R` file.
When students select a practice set from the RStudio Addin menu, an editor
window is opened with the prompts from the practice set.

Here the learner has responded to the prompts with the expected working 
code:

```
# pinfo201 / ps-1
#
# PS-Example: Example practice set
#    This file illustrates the essentials of specifying a practice set. It
#    show what kinds of prompts can be presented to #' students and what can be
#    tested.

# Practice set info ---- 
practice.begin("PS-Example", learner="[your name]", email="[your e-mail]")

# Key practice set variables (already initialized) ----
#   library(stats)
#   X <- c(1,2,3)

# Your 7 prompts: (a)-(g) ----

#                                         Note 01.
#    A practice set that illustrates the basics.


# a: Add ten, nine, and eight together. Assign the result to `sum1`. (Variable: sum1)
sum1 <- 10 + 9 + 8

# b: Create a variable `hometown` that stores the city in which you were born. (Variable: hometown)
hometown <- "St. Louis"

# c: Add 10 to each of the elements of vector `X`. (Variable: v1)
v1 <- X + 10

# d: Create 100 random numbers between 40000 and 50000. Round the numbers
#    to two decimal places. (Variable: salaries_2017)
salaries_2017 <- round(runif(100, 40000, 50000),2)

#                                         Note 02.
#    Working with functions.


# e: Write a function, named `what_is_pi`, which returns pi (3.1415). (Variable: what_is_pi)
what_is_pi <- function() {pi}

# f: Write a function, named `squared(x)`, which squares a number. (Variable: squared)
squared <- function(x) {
   t <- x^2
   return(t)
}

# g: Define a function, named `imperial_to_metric`, that takes in two arguments: a
#    number of feet and a number of inches. The function should return the
#    equivalent length in meters. (Variable: imperial_to_metric)
imperial_to_metric <- function(feet, inches) {
   total_inches <- feet * 12 + inches
   meters <- total_inches * 0.0254
   return(meters)
}

```

When a learner sources this file, the command `practice.begin()` associates the
code with the the practice set. When `practice.check()` (from the RStudio Addin
menu) is executed the student's code is checked and feedback is displayed in the
RStudio Viewer.


# A key assumption:

This package is oriented toward checking data values. Currently, it does not
model the structure of student's code nor its execution. For example, the
package cannot be used to evaluate the correctness of a task like this:

* _Use cat() to output two variables a and b._

It can, however, be used to evaluate the following:

* _Create a string from two variables, `a` and `b, and assign the string to `out`._

# Auto-grading

\code{pinfo201} can be used for auto-grading assistance. This function, for
example, will grade all assignments that are located in the directory,
\code{dir}:

\preformatted{
admin.grade(dir)
}

It does so by linking the code in the file to the problem set with the
\code{practice.begin()} command. Then, it parses and evaluates the R file.
Currently, the output is one \code{.html} file per \code{.R} file, with
each output file summarizing correct and incorrect solutions.

Other admin functions are used to assist in the development of practice
sets.


@section What can be evaluated?:
#'
Currently, \code{pinfo201} is able to check the following data types:
\itemize{
\item Special constants (NULL, NA, Inf, -Inf, and NaN)
\item Atomic vectors (scalars) (logical, numeric, integer, complex, character types)
\item Vectors greater than length 1 (comprising logical, numeric, integer, complex, and character types)
\item Lists
\item Data frames
\item Functions with zero, one, or two unnamed arguments (with the @check tag)
}
#'
@section Checking callbacks:
#'
At present, the framework can be extended with prompt-specific callback
functions for checking a learner's code, although this takes a good deal of
effort. Thus, if you need to test a function with 3 or more parameters or
test a data structure in some specific ways, you can write a checking
callback function. This is the current structure of the checking callbacks:
#'
#'\preformatted{
<funct_name>_Check <- function(internal_id, result) {
  learner_result <- eval_string_details(ps_get_assignment_var(internal_id))
  expected_result <- eval_string_details(ps_get_expected_answer(internal_id))
#'
  # Analyze learner's code (e.g., with an abstract syntax tree) and add feedback
  # to the result
  #     add.message(result, message)
#'
  if (identical(learner_result, expected_result)) {
    result <- result_update(result, internal_id, TRUE, result_good_msg(internal_id))
  } else {
    result <- result_update(result, internal_id, FALSE, result_error_msg(internal_id))
 }
 return(result)
 }}
#'
