---
title: "Introduction to pinfo201"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pinfo201}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pinfo201)
```

# Hello 

## Hello 
_pinfo201_ refers to _practice for INFO-201_. INFO-201 is an
introduction to data science at the University of Washington, a class originally
developed by Mike Freeman and Joel Ross. See:

* Freeman, M., & Ross, J. (2019). _[Programming Skills for Data Science: Start Writing Code to Wrangle, Analyze, and Visualize Data with R.](https://github.com/programming-for-data-science)_ Boston, MA: Addison-Wesley.



#' \code{pinfo201} is a package for practicing basic R programming. It is intended for
#' learners who are just beginning. In addition, \code{pinfo201} provides functions
#' for efficiently grading student work.
#'
#' @section Practice sets:
#'   The key idea is that an instructor creates a practice set, which comprises
#'   prompts, expected answers, and hints. Once \code{pinfo201} is installed,
#'   practice sets are loaded into RStudio and can be controlled with an Addins
#'   menu. The \code{pinfo201} menu includes these options:
#'   \itemize{
#'   \item 1. Begin Practice
#'   \item 2. Check Answers
#'   \item 3. Show Prompts
#'   \item 4. Show Answers}
#'
#'   Learners select from the available practice sets (\code{1. Begin
#'   Practice}). They write code for one or more of the prompts. When ready,
#'   learners check their code (\code{2. Check Answers}). If the learner's answer
#'   differs from the expected answer, hints, if available, are shown.
#'
#' @section Practice sets mark-up:
#'   Instructors write practice sets in \code{.R} files, with tagged comments. Here
#'   is a practice set, with illustrative prompts:

```
#'   \preformatted{
#'   #' @version ps-1
#'   #' @short PS-Example
#'   #' @title Example practice set
#'   #' @descr
#'   #' This file illustrates the essentials of specifying a practice set. It
#'   #' show what kinds of prompts can be presented to students and what can be
#'   #' tested.
#'   #' @end
#'   #' @initial-vars
#'   library(stats)
#'   X <- c(1,2,3)
#'   #' @end
#'
#'   #' @id -
#'   #' @msg
#'   #' A practice set that illustrates the basics.
#'   #' @end
#'
#'   #' @id a
#'   #' @msg Add ten, nine, and eight together. Assign the result to `sum1`.
#'   #' @code
#'   sum1 <- 10 + 9 + 8
#'   #' @end
#'   #' @hints
#'   #' Do you use the math plus operator (+)?
#'   #' Do you use the assignment operator (<-)?
#'   #' Is the expected variable name (sum1) used?
#'   #' @end
#'
#'   #' @id b
#'   #' @msg
#'   #' Create a variable `hometown` that stores the city in which you were born.
#'   #' @end
#'   #' @check list(re="^[a-zA-Z\\s\\.]*$")
#'   #' @code
#'   hometown <- "St. Louis"
#'   #' @end
#'
#'   #' @id c
#'   #' @msg Add 10 to each of the elements of vector `X`.
#'   #' @code
#'   v1 <- X + 10
#'   #' @end
#'
#'   #' @id d
#'   #' @msg
#'   #' Create 100 random numbers between 40000 and 50000. Round the numbers
#'   #' to two decimal places.
#'   #' @end
#'   #' @cp-var salaries_2017
#'   #' @code
#'   salaries_2017 <- round(runif(100, 40000, 50000),2)
#'   #' @end
#'
#'   #' @id -
#'   #' @msg Working with functions.
#'
#'   #' @id e
#'   #' @msg Write a function, named `what_is_pi`, which returns pi (3.1415).
#'   #' @var what_is_pi
#'   #' @code
#'   what_is_pi <- function() {pi}
#'   #' @end
#'
#'   #' @id f
#'   #' @msg Write a function, named `squared(x)`, which squares a number.
#'   #' @var squared
#'   #' @check list(arg1 = c(1, 2, 3, 0, -1, -2, -3, NA))
#'   #' @code
#'   squared <- function(x) {
#'     t <- x^2
#'     return(t)
#'   }
#'   #' @end
#'
#'   #' @id g
#'   #' @msg
#'   #' Define a function, named `imperial_to_metric`, that takes in two arguments: a
#'   #' number of feet and a number of inches. The function should return the
#'   #' equivalent length in meters.
#'   #' @end
#'   #' @check list(arg1 = c(4, 5, 100, 0, NA), arg2 = c(0, 1, 1.5, 12.0, 24))
#'   #' @code
#'   imperial_to_metric <- function(feet, inches) {
#'     total_inches <- feet * 12 + inches
#'     meters <- total_inches * 0.0254
#'     return(meters)
#'   }
#'   #' @end
#'   }
```
